### 2025-08-06 final-presentation/final.tex

\subsection*{Naturl deduction}
\begin{frame}{Natural Deduction}
\begin{itemize}
    \item \textbf{Natural Deduction} is a rule-based system for deriving conclusions from assumptions in logic.
    \item Instead of using exhaustive truth tables, proofs are built step-by-step using inference rules.
    \item Example: Proving from $A \land (A \rightarrow \bot)$ that $\bot$ (contradiction) can be derived.
    \item Basis for how proof assistants check the logical structure of proofs.
\end{itemize}
\end{frame}
\begin{frame}
    \begin{figure}[h]
  \centering
  \includegraphics[width=0.8\textwidth]{nd.jpeg}
\end{figure}
\end{frame}

\subsection*{Ins}
\begin{frame}{Intuitionistic Logic}
\begin{itemize}
    \item \textbf{Intuitionistic Logic} Also called Constructive Logic, reflects principles of constructive mathematics, where a statement is only true if a proof can be constructed.
    \item Omits some classical logic rules, such as the Law of Excluded Middle.
    \item Stronger requirement: to prove existence, a method or algorithm must be given.
    \item Proof assistants leverage this constructive approach for digital verification.
\end{itemize}
\end{frame}

@@

\end{figure}
Inference Rules for Intuitionistic Logic
\end{frame}
\subsection*{$\lambda$-Calculus}
\begin{frame}{$\lambda$-Calculus and Type Theory}
\begin{itemize}
    \item \textbf{$\lambda$-Calculus}: A foundational system for defining and applying functions using abstraction and application.
    \item \textbf{Type Theory}: Assigns types to every term; ensures correctness of operations.
         \item \emph{Dependent types} allow types to depend on values, expressing complex logical properties.

    \item \textbf{Curry--Howard Correspondence}: 
    \begin{itemize}
        \item Propositions $\leftrightarrow$ Types
        \item Proofs $\leftrightarrow$ Programs
    \end{itemize}
\end{itemize}
\end{frame}
\section{Architecture of proof assistant}
\begin{frame}{Architecture of a Proof Assistant}
\begin{itemize}
    \item \textbf{Kernel}: Minimal, trustworthy codebase enforcing logical rules and validating proofs.
    \item \textbf{Tactic Engine}: Helps build and automate proofs step by step.
    \item \textbf{Formal Proof Language}: Rigorously expresses definitions, statements, and proofs.
    \item \textbf{Libraries}: Collections of verified mathematical foundations for reuse.
    \item \textbf{User Interface}: IDEs and plugins for interactive, efficient proof development.
\end{itemize}
\end{frame}
\subsection*{Kernel}
\begin{frame}{Kernel: The Trusted Core}
\begin{itemize}
    \item The \textbf{kernel} is the minimal and most critical part of a proof assistant.
    \item It enforces the logical rules of the underlying formal system (e.g., type theory).
    \item Responsible for \textbf{validating every proof step} to guarantee correctness.
    \item Ensures \textbf{soundness and trustworthiness}; the rest of the system depends on its integrity.
    \item Typically very small and rigorously tested or formally verified to avoid bugs.
    \item Example: Agda’s kernel is written in Haskell and integrates normalization to check definitional equality.
\end{itemize}
\end{frame}
\subsection*{Tactic Engine}

@@

\section{Formalization Of Some proofs}
\subsection*{}

>>>>

\begin{frame}{Natural Deduction}
\begin{itemize}
    \item \textbf{Natural Deduction} is a rule-based system for deriving conclusions from assumptions in logic.
    \item Instead of using exhaustive truth tables, proofs are built step-by-step using inference rules.
    \item Example: Proving from $A \land (A \rightarrow \bot)$ that $\bot$ (contradiction) can be derived.
\end{itemize}
\end{frame}
\begin{frame}
    \begin{figure}[h]
  \centering
  \includegraphics[width=0.8\textwidth]{nd.jpeg}
\end{figure}
\end{frame}

\begin{frame}{Intuitionistic Logic}
\begin{itemize}
    \item \textbf{Intuitionistic Logic} Also called Constructive Logic, reflects principles of constructive mathematics, where a statement is only true if a proof can be constructed.
    \item Omits some classical logic rules, such as the Law of Excluded Middle.
    \item Stronger requirement: to prove existence, a method or algorithm must be given.
    \item Proof assistants leverage this constructive approach for digital verification.
\end{itemize}
\end{frame}

@@

\end{figure}
Inference Rules for Intuitionistic Logic
\end{frame}
\begin{frame}{Lambda Calculus}
\textbf{What is Lambda Calculus?}
\begin{itemize}
  \item A formal model of computation by Alonzo Church
  \item Lambda Terms:
    \begin{itemize}
        \item Variables: $x$, $y$, $z$
        \item Abstraction: $\lambda x.E$
        \item Application: $(\lambda x.E)F$
    \end{itemize}
\end{itemize}
\end{frame}
\begin{frame}
\textbf{Examples}
\begin{itemize}
    \item $\lambda x.x^2$ is a function
    \item $(\lambda x.x^2)(3) \rightarrow 9$
\end{itemize}

\begin{frame}{Type Theory \& Simply Typed Lambda Calculus}
\textbf{Type Theory Basics}
\begin{itemize}
  \item Assigns types to terms: $1 : \mathbb{N}$, $+ : \mathbb{N} \times \mathbb{N} \to \mathbb{N}$
  \item Typing is decidable
\end{itemize}

\textbf{Type Categories}
\begin{itemize}
  \item Base Types: $\mathbb{N}$, Bool, $\bot$
  \item Arrow Types: $f : A \to B$
  \item Product Types: $\langle a, b \rangle : A \times B$
  \item Sum Types: $a : A + B$
\end{itemize}
\end{frame}
\begin{frame}
   \textbf{Typed Lambda Calculus}
\begin{itemize}
  \item $(\lambda x : \mathbb{N}.x^2) : \mathbb{N} \to \mathbb{N}$
\end{itemize}
 \textbf{Dependent Types}
\begin{itemize}
  \item Types depend on values: $Vec(n)$
  \item Indexed types, predicate representation
  \item $Vec : \mathbb{N} \to Type$
\end{itemize}

\begin{frame}{ Curry–Howard Correspondence}
    \begin{itemize}
  \item sometimes referred as Curry Howard Isomorphism
  \item A connection between logic, computation, and type theory
  \item Also known as the \textit{proofs-as-programs} and \textit{propositions-as-types} principle
\end{itemize}

\textbf{Core Idea}
\begin{itemize}
  \item A \textbf{proposition} corresponds to a \textbf{type}
  \item A \textbf{proof} of the proposition is a \textbf{program} (term) of that type
  \item Proof checking is equivalent to type checking
  \item Proof normalization corresponds to program evaluation
\end{itemize}
\end{frame}

\begin{frame}
\textbf{Curry–Howard Correspondence}
\begin{tabular}{|c|c|c|}
\hline
Logic & Type Theory & Programming \\
\hline
Proposition & Type & - \\
Proof & Term & Program \\
Implication $A \to B$ & $A \to B$ & $\lambda$-abstraction \\
Conjunction $A \land B$ & $A \times B$ & Pair \\
Disjunction $A \lor B$ & $A + B$ & Tagged union \\
Falsehood $\bot$ & Empty Type & No term \\
Universal $\forall x.A(x)$ & $\Pi x:A.B(x)$ & Function over types \\
Existential $\exists x.A(x)$ & $\Sigma x:A.B(x)$ & Dependent pair \\
\hline
\end{tabular}

    \begin{frame}{Martin-Löf Type Theory}
\textbf{What is it?}
\begin{itemize}
  \item A formal system for constructive mathematics
  \item Also known as Intuitionistic Type Theory
  \item Backbone of modern Proof Assistants
\end{itemize}

\textbf{Core Types}
\begin{itemize}
  \item \textbf{} ($\Pi$-type): Dependent function type, $\forall x:A.B(x)$
  \item \textbf{} ($\Sigma$-type): Dependent sum type, $\exists x:A.B(x)$
  \item \textbf{Identity Type}: Internal equality between terms
\end{itemize}

\begin{frame}
  \textbf{Type Universes}
\begin{itemize}
  \item Types have types: $1 : \mathbb{N}, \mathbb{N} : Type, Type : \cdots$
  \item Not like sets within sets — avoids paradoxes
  \item Enables reasoning and abstraction over types themselves

\end{itemize}  
\end{frame}

\section{Architecture of proof assistant}
\begin{frame}{Architecture of a Proof Assistant}
\begin{itemize}
    \item \textbf{Kernel}: Minimal, trustworthy codebase enforcing logical rules and validating proofs.
    \item \textbf{Tactic Engine}: Helps build and automate proofs step by step.
    \item \textbf{Formal Proof Language}: Rigorously expresses definitions, statements, and proofs.
    \item \textbf{Libraries}: Collections of verified mathematical foundations for reuse.
    \item \textbf{User Interface}: IDEs and plugins for interactive, efficient proof development.
\end{itemize}
\end{frame}
\subsection*{Kernel}
\begin{frame}{Kernel: The Trusted Core}
\begin{itemize}
    \item The \textbf{kernel} is the minimal and most critical part of a proof assistant.
    \item It enforces the logical rules of the underlying formal system (e.g., type theory).
    \item Responsible for \textbf{validating every proof step} to guarantee correctness.
    \item Ensures \textbf{soundness and trustworthiness}; the rest of the system depends on its integrity.
    \item Typically very small and rigorously tested or formally verified to avoid bugs.
    \item Example: Agda’s kernel is written in Haskell and integrates normalization to check definitional equality.
\end{itemize}
\end{frame}
\subsection*{Tactic Engine}

@@

\section{Formalization Of Some proofs}
\subsection*{Defining Natural Numbers} 
    \begin{frame}[fragile]
        \frametitle{Eg: Defining Natural Numbers}
        \begin{verbatim}
           data N : Set where 
                Zero : N 
                suc : N -> N
        \end{verbatim}
    \end{frame}
